Ontology Matching with word2vec

### 2016-05-15 Paper/Draft.tex
Feedback up to page 16

 \subsection{Other related research}
 \subsubsection{Word sense representation} \label{w2v}
 Skip-gram models have been widely used to represent words as vectors. This is useful as it is easy to calculate the distance between two vectors, which is a useful property for ontology matching. This vector distance has been shown to be related to the semantic similarity between words, which is very useful for ontology alignment, as parts of one ontology have to be aligned with parts of another ontology depending on their semantic relation.

 Recently, an effective and efficient skip-gram model has been developed, called word2vec, which uses a number of extensions over previous methods that enable it to efficiently learn an effective representation/cite{w2v}.
 Word2vec learns this representation by trying to predict the context of a word. First the vocabulary is one-hot encoded. Then the model learns to predict a context vector that minimizes the distance to the context one-hot encoded vectors. The model is a neural network with one hidden layer. Since the input is one-hot encoded, every word in the vocabulary is represented by one row in the weight matrix between the input layer and hidden layer. This row is the vector representation of the word.

 Word2vec has been expanded in a number of different ways. For example, word order can be preserved, leading to a similarity measure that is closer to syntax, as syntax defines which words go where in a sentence. This preservation was done by having multiple output layers, one for every word in the context. The research shows that the input does not have to be a bag-of-words representation, which will be useful later on\cite{ling2015}.
 Another interesting development is the multi-sense word2vec, which allows for multiple vector representations per word, depending on the number of different definitions a word has. For example, the word bank represents both the monetary institute and the riverside, which would both have a different representation in this extension. It can differentiate between two different meanings based on the context. This will help with resolving ambiguity, a very important factor in ontology alignment\cite{multisense}.
 \subsection{Used technology}
 \subsubsection{AgreementMaker} \label{agreementmaker}
 AgreementMaker is an ontology matching system that obtained the highest F-measure in 6 of the 7 ontology matching tracks of OAEI 2015 \cite{OAEI2015}. The system has also been updated the most recent at the time of writing\cite{agreementmaker}.

 The system uses an extensible framework which allows its users to add new modules to the system. This is very useful for research as it allows the researcher to compare different modules by swapping in just those modules in the framework. The module-swapping technique will also be used in this work, but more of that will be discussed in \ref{Evaluation methods}.

 \subsubsection{Factorie} \label{factorie}
 "FACTORIE [is a] a toolkit for probabilistic modeling based on imperatively-defined factor graphs" \cite{git-factorie}. As a toolkit it contains many different parts that can be used together. One example is the NLP package in the toolkit, which provides many algorithms to process text. One of those algorithms is the distributed representation of words algorithm by \cite{w2v}.

@@

 \subsection{Problem}
 We have seen that when information from multiple domains needs to be integrated so that it is possible to reason over them together, the ontologies of those domains need to be aligned or merged. % Waar?
  Those ontologies contain rules about concepts and concept classes in the shape of triplets, with two concepts and a relation. For example, the rule {\tt Professors are humans} contains the concepts \emph{Professor} and \emph{human} and relationship \emph{being}.

@@

 Lastly, there is the problem of different types of alignment relations. For example, one concept can be a \emph{part of} another concept, or an example of one. These different relations all require a different approach and come with their own sets of problems.

@@

 \begin{center}
 \textit{
 Can distributed representations perform well in a domain it was not designed for?
 }
 \end{center}

 \subsection{Hypotheses}
 Since multiple implementations of the algorithm are considered, every hypothesis that follows is considered for every implementation.

 \begin{center}
 \textit{This specific matcher that uses a distributed representation of concepts can find relationships between concepts in ontologies better than existing terminological matchers.}
 \end{center}

@@

 \begin{center} 
 \textit{The distributed representations matcher can use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers.}
 \end{center}
 \begin{center}
 \textit{The distributed representations matcher can perform well in a domain it was not designed for.}
 \end{center}

 The following hypotheses are true if for at least one implementation the corresponding hypothesis is true.

 \begin{center}  
 \textit{A distributed representations of concepts matcher can find relationships between concepts in ontologies better than existing terminological matchers.}
 \end{center} 
 \begin{center}
 \textit{A distributed representations matcher can use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers.}
 \end{center}
 \begin{center}
 \textit{A distributed representations matcher can perform well in a domain it was not designed for.}
 \end{center}

@@

 Since distributed representation algorithms can be trained unsupervised on new domains, they can be applied to those new domains without effort from humans, unlike current matching algorithms. This is also an improvement for the semantic web.

 However, ontology matching is also used outside of the semantic web. It is already in use for companies that want to merge knowledge from other companies, for example. If the proposed matching algorithm is better than the current algorithms, or at least improves the performance when used in ensemble with the current algorithms, it will be useful in any situation, not just on the Semantic Web.

 \newpage
 \section{Method}
 In this section, I will describe the method that I developed. Firstly, I will describe the general idea. In the second section, I will give a theoretical justification for why this method should work and how it works conceptually. After that I will describe the different ways that the input graphs can be used to train the word2vec model and create the alignment, some of which I have discarded on theoretical grounds as I will explain in the relevant section. Lastly, I will talk about the implementation process.
 \subsection{Idea}
 As explained earlier, word2vec is a model that learns two things: implicitly it learns the semantic vector representation of words, and explicitly it learns to predict words from their context or vice versa. I will use both of these pieces of information in our alignment algorithm.
 The semantic representation will be used for finding nodes in the two ontology graphs that are similar, i.e. close in the vector space. This step is relatively straightforward if proper vector representations are found. However, this may be hard due to a number of problems which I have listed below.

@@

 To solve these problems I will adapt word2vec to graphs and extend it with multi-sense embeddings. Also hot-starting is considered as an improvement.

@@

 In this section, I will discuss the different implementations that were made of the algorithm. All implementations use the same basic algorithm in different ways by taking into account the context in different ways. All of these implementations were tested and the results can be found in \ref{results}. Some implementations are extensions of other implementations, but all implementations are mentioned for completeness and to enable explanation of the source of improvements.

@@

 \subsection{Implementation}
 The two foundations the algorithm has been build and tested on are Factorie, described in \ref{factorie}, and AgreementMaker, described in \ref{agreementmaker}, respectively. Earlier work on a forked branch of factorie was done by \cite{multisensecode} to implement multi-sense word2vec.

@@

 Word2vec is an unsupervised method that maps words to a vector space that has been shown to contain semantic properties. %cite
 It can be trained on large corpora quickly and the resulting vectors can easily be used to measure the distance between words. All these properties make it an excellent candidate for a concept distance measure to use for ontology matching. As an added benefit, it can also take into account the context of a concept, which may be able to improve its performance over a matcher that only uses the label of a concept.
 \subsection{Theory}

>>>>

 \subsection{Word sense representation} \label{w2v}
 Skip-gram models have been widely used to represent words as vectors. This is useful as it is easy to calculate the distance between two vectors, which is a useful property for ontology matching. This vector distance has been shown to be related to the semantic similarity between words, which is very useful for ontology alignment, as parts of one ontology have to be aligned with parts of another ontology depending on their semantic relation.

 Recently, an effective and efficient skip-gram model has been developed, called \textf{word2vec}, which uses a number of extensions over previous methods that enable it to efficiently learn an effective representation\cite{w2v}.
 \textf{Word2vec} learns this representation by trying to predict the context of a word. First the vocabulary is one-hot encoded. Then the model learns to predict a context vector that minimizes the distance to the context one-hot encoded vectors. The model is a neural network with one hidden layer. Since the input is one-hot encoded, every word in the vocabulary is represented by one row in the weight matrix between the input layer and hidden layer. This row is the vector representation of the word.

 \textf{Word2vec} has been expanded in a number of different ways. For example, word order can be preserved, leading to a similarity measure that is closer to syntax, as syntax defines which words go where in a sentence. This preservation was done by having multiple output layers, one for every word in the context. The research shows that the input does not have to be a bag-of-words representation, which will be useful later on\cite{ling2015}.
 Another interesting development is the multi-sense \textf{Word2vec}, which allows for multiple vector representations per word, depending on the number of different definitions a word has. For example, the word bank represents both the monetary institute and the riverside, which would both have a different representation in this extension. It can differentiate between two different meanings based on the context. This will help with resolving ambiguity, a very important factor in ontology alignment\cite{multisense}.
 \subsection{Used technology}
 %TODO Guide user from problem to research question. See Arjen's notes 1. 
 \subsubsection{Factorie} \label{factorie} %TODO rewrite
 "FACTORIE [is a] a toolkit for probabilistic modeling based on imperatively-defined factor graphs" \cite{git-factorie}. As a toolkit it contains many different parts that can be used together. One example is the NLP package in the toolkit, which provides many algorithms to process text. One of those algorithms is the distributed representation of words algorithm by \cite{w2v}.

@@

 \subsection{Problem}
 We have seen that when information from multiple domains needs to be integrated so that it is possible to reason over them together, the ontologies of those domains need to be aligned or merged. % Waar?
  Those ontologies contain rules about concepts and concept classes in the shape of triplets, with two concepts and a relation. For example, the rule {\tt Professors are humans} contains the concepts \emph{Professor} and \emph{human} and an \emph{is-a}-relationship.

@@

 Lastly, there is the problem of different types of alignment relations. For example, one concept can be a \emph{part of} another concept, or an example of one. These different relations all require a different approach and come with their own sets of problems. %Meronymy?

@@

 \begin{center}
 \textit{
 Can distributed representations perform well in a domain it was not trained for?
 }
 \end{center}

 \subsection{Hypotheses} 
 \begin{center}
 \textit{This specific matcher that uses a distributed representation of concepts can find relationships between concepts in ontologies better than existing terminological matchers.}
 \end{center}

@@

 \begin{center} 
 \textit{The distributed representations matcher can use the context of a concept to find relationships between concepts in ontologies better than existing structural matchers.}
 \end{center}
 \begin{center}
 \textit{The distributed representations matcher can perform well in a domain it was not trained for.}
 \end{center}

 These hypotheses are true for distributed representations matcher if they are true for one distributed representation matcher. Therefore to test the hypotheses, I must test them for every distinct implementation of the matcher.

@@

 Since distributed representation algorithms can be trained unsupervised on new domains, they can be applied to those new domains without effort from humans, as opposed to for example the string edit distance. This is also an improvement for the semantic web.

 However, ontology matching is also used outside of the semantic web. It is already in use for companies that want to merge knowledge from other companies, for example. If the proposed matching algorithm is better than the current algorithms, or at least improves the performance when used in ensemble with the current algorithms, it will be useful in these situations, not just on the Semantic Web.

 \newpage
 \section{Method}
 This Section will describe the method that I developed. Firstly, I will describe the general idea. In the second section, I will give a theoretical justification for why this method should work and how it works conceptually. After that I will describe the different ways that the input graphs can be used to train the \textf{Word2vec} model and create the alignment, some of which I have discarded on theoretical grounds as I will explain in the relevant section. Lastly, I will talk about the implementation process.
 \subsection{Idea}
 As explained earlier, \textf{word2vec} is a model that learns two things: implicitly it learns the semantic vector representation of words, and explicitly it learns to predict words from their context or vice versa. I will use both of these pieces of information in our alignment algorithm.
 The semantic representation will be used for finding nodes in the two ontology graphs that are similar, i.e. close in the vector space. This step is relatively straightforward if proper vector representations are found. However, this may be hard due to a number of problems which I have listed below.

@@

 To solve these problems I will adapt \textf{Word2vec} to graphs and extend it with multi-sense embeddings. Also hot-starting is considered as an improvement. %TODO this paragraph needs to be explained

@@

 In this section, I will discuss the different implementations that were made of the algorithm. All implementations use the same basic algorithm in different ways by taking into account the context in different ways. All of these implementations were tested and the results can be found in Section \ref{results}. Some implementations extend others, but all implementations are covered to ensure completeness. This way, all improvements are recorded precisely.

@@

 \subsection{Implementation}
 The two foundations the algorithm has been build and tested on are Factorie, described in \ref{factorie}, and AgreementMaker, described in \ref{agreementmaker}, respectively. Earlier work on a forked branch of factorie was done by \cite{multisensecode} to implement multi-sense \textf{Word2vec}.

@@

 \textf{Word2vec} is an unsupervised method that maps words to a vector space that has been shown to contain semantic properties. %cite
 It can be trained on large corpora quickly and the resulting vectors can easily be used to measure the distance between words. All these properties make it an excellent candidate for a concept distance measure to use for ontology matching. As an added benefit, it can also take into account the context of a concept, which may be able to improve its performance over a matcher that only uses the label of a concept.
 \subsection{Theory}
